### 2018-07-17 notes/nbeSum.tex
ITC lecture: reflect/reify for implication

\nonstopmode
\documentclass[a4paper]{article}
\usepackage{natbib}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{amsmath}
%\usepackage{dsfont}

\input{macros}

\newtheorem{theorem}{Theorem}
%\newtheorem*{theorem*}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}

%\newcommand{\qed}{\hfill\ensuremath{\Box}}
% from llncs.cls
%\def\squareforqed{\hbox{\rlap{$\sqcap$}$\sqcup$}}
\def\squareforqed{\ensuremath{\Box}}
\def\qed{\ifmmode\squareforqed\else{\unskip\nobreak\hfil
\penalty50\hskip1em\null\nobreak\hfil\squareforqed
\parfillskip=0pt\finalhyphendemerits=0\endgraf}\fi}

\newenvironment{proof}[1][]{\noindent\ifthenelse{\equal{#1}{}}{{\it
      Proof.}}{{\it Proof #1.}}\hspace{2ex}}{\qed\bigskip}
\newenvironment{proof*}[1][]{\noindent\ifthenelse{\equal{#1}{}}{{\it
      Proof.}}{{\it Proof #1.}}\hspace{2ex}}{\bigskip}

\newcommand{\wk}{\mathord{\uparrow}}
\newcommand{\WW}{\mathbb{W}}
\newcommand{\tAr}{\mathsf{Ar}}
\newcommand{\Ar}[1]{\tAr(#1)}

\title{Normalization by Evaluation for \\ Intuitionistic Propositional Logic}
\author{Andreas Abel}
\date{July 2018}

\begin{document}
\maketitle

\section{Intuitionistic Propositional Logic (IPL)}

\para{Formulas} and hypotheses lists (contexts).
\[
\begin{array}{lllrll}
  \Atom & \ni & P,Q & & & \mbox{atomic propositions} \\
  \Form & \ni & A,B,C,\dots
     & ::= & \top \mid A \land B \mid A \imp B
       & \mbox{negative formulas} \\
     &&& \mid & \bot \mid A \lor B \mid P
       & \mbox{positive formulas} \\
\\[1ex]
  \Cxt & \ni & \Gamma,\Delta & ::= & \cempty & \mbox{empty context}
\\ &&& \mid & \Gamma.A & \mbox{context extension}
\end{array}
\]
Indices $x$ into the context $\Gamma$ of hypothesis, locating hypothesis
$A$, are written $x : \Hyp[\Gamma]A$ and defined inductively by the
following constructors:
\[
  \ru{}{\ttop : \Hyp[\Gamma.A]A}
\qquad
  \ru{x : \Hyp[\Gamma]A
    }{\pop[B] x : \Hyp[\Gamma.B]A}
\]
Deriviations $t$ of the truth of a formula $A$ under assumptions
$\Gamma$ (judgement $\Gamma \der A$)
are written $t : \Tm[\Gamma]A$ and defined inductively as follows:

\paradot{Implication and hypotheses}
\begin{gather*}
  \ru{x : \Hyp[\Gamma]A}{\hyp x : \TmG A}
\qquad
  \ru{t : \Tm[\Gamma.A]B}{\impI t : \TmG{A \imp B}}
\qquad
  \ru{t : \TmG{A \imp B} \qquad
      u : \TmG A
    }{\impE t u : \TmG B}
\end{gather*}

\paradot{Truth and conjunction}
\begin{gather*}
  \ru{}{\trueI : \TmG\top}
\qquad
  \ru{t : \TmG A \qquad
      u : \TmG B
    }{\andI t u : \TmG{A \land B}
    }
\\[1ex]
  \ru{t : \TmG{A \land B}
    }{\andE1 t : \TmG A
    }
\qquad
  \ru{t : \TmG{A \land B}
    }{\andE2 t : \TmG B
    }
\end{gather*}

\paradot{Absurdity and disjunction}
\begin{gather*}
  \ru{t : \TmG A
    }{\orI1 t : \TmG {A \lor B}
    }
\qquad
  \ru{t : \TmG B
    }{\orI2 t : \TmG {A \lor B}
    }
\\[1ex]
  \ru{t : \TmG \bot
    }{\falseE t : \TmG C}
\qquad
  \ru{t : \TmG {A \lor B} \qquad
      u : \Tm[\Gamma.A]C  \qquad
      v : \Tm[\Gamma.B]C
    }{\orE t u v : \TmG C}
\end{gather*}

\section{Subformula property and normal derivations}

For every derivable judgement $\Gamma \der A$ there are infinitely
many derivations $t : \TmG A$, since arbitrary detours are allowed.
For instance to prove $\top$ we can also proceed very indirectly by
introducing a hypothesis $\top$ and then eliminating it by proof
$\trueI$:
\[\impEP{\impIP{\hyp{\ttop}}}{\trueI} : \Tm[\cempty]\top.\]
A sensible restriction is that the proof rules guarantee that only
\emph{subformulas} of $A$ and the hypotheses in $\Gamma$ are mentioned when
deriving $\Gamma \der A$.

The following indexed grammar for \emph{normal} derivations $\NfG A$ of
judgement $\Gamma \der A$ ensures the subformula property.  It is
defined mutually with a grammar $\NeG A$ of \emph{neutral} derivations.
Neutral derivations are ``straightline'' consequences of hypotheses
(without case distinction or absurdity elimination).

\paradot{Implication and hypotheses}
\begin{gather*}
  \ru{x : \Hyp[\Gamma]A}{\hyp x : \NeG A}
\qquad
  \ru{t : \Nf[\Gamma.A]B}{\impI t : \NfG{A \imp B}}
\qquad
  \ru{t : \NeG{A \imp B} \qquad
      u : \NfG A
    }{\impE t u : \NeG B}
\end{gather*}

\paradot{Truth and conjunction}
\begin{gather*}
  \ru{}{\trueI : \NfG\top}
\qquad
  \ru{t : \NfG A \qquad
      u : \NfG B
    }{\andI t u : \NfG{A \land B}
    }
\\[1ex]
  \ru{t : \NeG{A \land B}
    }{\andE1 t : \NeG A
    }
\qquad
  \ru{t : \NeG{A \land B}
    }{\andE2 t : \NeG B
    }
\end{gather*}

\paradot{Absurdity and disjunction}
\begin{gather*}
  \ru{t : \NfG A
    }{\orI1 t : \NfG {A \lor B}
    }
\qquad
  \ru{t : \NfG B
    }{\orI2 t : \NfG {A \lor B}
    }
\\[1ex]
  \ru{t : \NeG \bot
    }{\falseE t : \NfG C}
\qquad
  \ru{t : \NeG {A \lor B} \qquad
      u : \Nf[\Gamma.A]C  \qquad
      v : \Nf[\Gamma.B]C
    }{\orE t u v : \NfG C}
\end{gather*}

\paradot{Embedding}
\[
  \ru{t : \NeG P
    }{\neu t : \NfG P}
\]

The introduction rules generate normal forms ($\tNf$s) from $\tNf$s.  The elimination
rules for negative formulas generate neutrals ($\tNe$s) from $\tNe$s, where
$\timpE$ requires a side argument in normal form.

Some care is needed with  the elimination rules for
positive formulas which are $\tfalseE$ and $\torE$.
These prove an arbitrary formula $C$, which, for the sake of the
subformula property, should not be subject to
further elimination but directly prove the goal.
Thus, in addition to restriction the elimination to neutrals
(perform case distinction only on neutrals),
we require them to produce a normal form
in case of $\torE$ from normal branches.

Any neutral derivation at atomic proposition $P$ is considered normal.
The restriction on atoms is not needed for the subformula property,
but forces derivations to be $\eta$-long.
We could drop the restriction, but even with this restriction, the
calculus is complete, meaning that for every derivation ($\tTm$) there
exists a normal derivation ($\tNf$).  This statement is called
\emph{normalization} and we will prove it constructively in the
following.

\section{Categories and Presheaves}

We define order-preserving embeddings (OPEs) $\tau : (\Delta \leq \Gamma)$ of a
context $\Gamma$ into a larger context $\Delta$ inductively by the
following constructors.
\begin{gather*}
  \ru{}{\id[\Gamma] : \Gamma \leq \Gamma}
\qquad
  \ru{\tau : \Delta \leq \Gamma
    }{\weak[A] \tau : \Delta.A \leq \Gamma}
\qquad
  \ru{\tau : \Delta \leq \Gamma
    }{\lift[A] \tau : \Delta.A \leq \Gamma.A}
\end{gather*}
For example $\weak [D] (\lift[C] (\weak[B] \id[A])) : A.B.C.D \leq
A.C$.  OPEs let us introduce extra, unused hypotheses into the
context.  Derivations of $\Gamma \der A$ can be weakened to
derivations $\Delta \der A$ for $\Delta \leq \Gamma$.

OPEs are closed under composition:  If $\tau : \Delta \leq \Gamma$ and
$\tau' : \Phi \leq \Delta$ then $\tau \comp \tau' : \Phi \leq \Delta$.
Let $\OPE$ be the category of contexts (as objects) and OPEs (as morphisms).
\begin{exercise}
  Define composition of OPEs by induction and show that $\OPE$ is
  indeed a category.
\end{exercise}

The context-indexed sets $\Hyp A$, $\Tm A$, $\Nf A$, and $\Ne A$ are
closed under weakening. \Eg{} $\tHyp$: for each $\tau : \Delta \leq \Gamma$
we have a function
\[
  \Hyp[\tau]A : \Hyp[\Gamma] A \to \Hyp[\Delta] A,
\]
\ie, a morphism in the category $\SET$ of sets and functions.
The collection $\Hyp A : (\tau : \Delta \leq \Gamma) \to
(\Hyp[\Gamma]A \to \Hyp[\Delta]A)$ of these function constitutes a
contravariant functor $\Hyp A : \OPE \to \SET$ mapping OPEs to
functions.  In other words, $\Hyp A$ is a \emph{presheaf} over $\OPE$
for each $A$, and so are $\Tm A$, $\Nf A$, and $\Ne A$.

\begin{exercise}
  Prove that $\Hyp A$, $\Tm A$, $\Nf A$, and $\Ne A$ are presheaves
  over $\OPE$ by defining the map functions and proving the functor laws!
\end{exercise}

A function $f$ between presheaves $\A$ and $\B$ is defined pointwise;
we write $f : \A \todot \B$ for the $\Cxt$-indexed collection
\[
  f_\Gamma : \A_\Gamma \to \B_\Gamma
\]
of functions (morphisms in $\SET$). Presheaves and functions between
them make a category by virtue of pointwise identity and composition.
However, typically the functions we consider are \emph{natural} in the
context index, \ie, they are \emph{natural transformations} between
presheaves, which means that they commute with OPEs.  Given $\tau :
\Delta \leq \Gamma$ and $f : \A \todot \B$, naturality means
\[
  f_\Delta \comp \A_\tau = \B_\tau \comp f_\Gamma .
\]
In words, first weakening with $\tau$ and then applying $f$ has the
same effect as applying $f$ first and perform the weakening later.
This guarantees some form of \emph{parametricity} of $f$ in the
context, in particular, $f$ cannot make decisions based on the length
of the context or its precise contents.

Presheaves over $\OPE$ and natural transformations form a category
$\PSh$, traditionally called $\widehat{\OPE}$.

\begin{exercise}
  Prove that $\PSh$ is indeed a category.
\end{exercise}

\begin{exercise}
  Prove that the following functions are morphisms in $\PSh$:
\[
\begin{array}{lllll}
  \thyp & : & \Hyp A & \todot & \Ne A \\
  \tandE1 & : & \Ne{A \land B} & \todot & \Ne A \\
  \torI1  & : & \Nf A & \todot & \Nf{A \lor B} \\
  \tfalseE & : & \Ne \bot & \todot & \Nf C \\
\end{array}
\]
  Find other examples of presheaf morphisms.
\end{exercise}

\section{NbE for the negative fragment}

We wish to define a normalization function for derivations of any
formula $A$
\[
  \tnorm^A : \Tm A \todot \Nf A .
\]
Normalization by evaluation (NbE) achieves normalization by first
evaluating terms, obtaining values and functions that compute by means
of the computation in the meta language, i.e., in $\SET$.  As a second
step, these values are reified back into syntax, yielding a normal
form.  Technically, we define a suitable presheaf $\den A$ for each
formula $A$ and two functions
\[
\begin{array}{lllll}
  \teval^A & : & \Tm A  & \todot & \den A \\
  \treify^A  & : & \den A & \todot & \Nf A \\
\end{array}
\]
which compose to $\tnorm$.

\subsection{Semantics, reflection and reification}

For now, let us focus on the negative
formulas $\top$, $A \land B$, $A \imp B$ and atoms $P$.
\[
\begin{array}{lll@{\qquad}lll}
  \den P & = & \Nf P
    & \reify[\Gamma] P t & = & t \\
  \den \top & = & \hatone
    & \reify[\Gamma] \top \_ & = & \trueI \\
  \den{A \land B} & = & \den A \hattimes \den B
    & \reify[\Gamma]{A \land B} {(a,b)} & = & \andIPP {\reify[\Gamma] A a} {\reify[\Gamma] B b} \\
\end{array}
\]
Herein, we use the unit presheaf $\hatone_\Gamma = 1$ where $1$ is the
unit set, and the pointwise product of presheaves
$(\A \hattimes \B)_\Gamma = \A_\Gamma \times \B_\Gamma$.  We note that
by these definitions, the category $\PSh$ has products, with the
pointwise terminal morphisms:
\[
\begin{array}{lll@{\qquad}lll}
   ! & : & \A \todot \hatone &
     [\_,\_] & : & (f : \C \todot \A) (g : \C \todot \B) \to (\C
                   \todot \A \hattimes \B) \\
   !_\Gamma(a) & = & ()       &
     [f,g]_\Gamma(c) & = & (f(c),g(c)) \\
\end{array}
\]

The interpretation $\den{A \imp B}$ of implication will follow the
Brouwer-Heyting-Kolmogorov (BHK) interpretation of intuitionistic
logic:
\begin{center}\itshape
  A proof of $A \imp B$ is a method turning any proof of $A$ into a
  proof of $B$.
\end{center}
However, the direct lifting of the functions space $\den{A \imp
  B}_\Gamma = \den A_\Gamma \to \den B_\Gamma$ does not work.  It does
not give a presheaf because of the contravariant occurrence of
presheaf $\den A$.  Instead, we force monotonicity by quantifying over
all OPEs of $\Gamma$:
\[
  \den{A \imp B}_\Gamma = \prod_{\tau : \Delta \leq \Gamma} (\den A
  _\Delta \to \den B _\Delta)
\]
Reification is also challenging.  Let $f : \den{A \imp B}_\Gamma$ and
attempt
%$\tau : \Delta \leq \Gamma$
\[
  \reify[\Gamma]{A \imp B} f %\;\Delta\;\tau =
    = \impIP {\reify[\Gamma.A]B{(f\;(\weak[A]{\id[\Gamma]})}\;\mbox{?})}
\]
where hole ``?'' should be filled with an element of $\den B_{\Gamma.A}$
which intuitively should represent the new hypothesis $\top :
\Hyp[\Gamma.A]A$ introduced by $\timpI$.  Thus we need a natural
transformation $\Hyp A \todot \den A$ which \emph{reflects} variables
into the semantics.  In order to define it by
induction on $A$, we need to generalize it to
\[
  \treflect^A : \Ne A \todot \den A
  .
\]
Let us make sure we can define it at atoms and conjunctions:
\[
\begin{array}{lllll}
  \treflect^P & = & \tne & : & \Ne P \todot \Nf P \\
  \treflect^\top & = & !  & : & \Ne \top \todot \hatone \\
  \treflect^{A \land B} & = & [\, \treflect^A \comp \andE1  , \;
                             \treflect^B \comp \andE2 ]
    & : & \Ne{A \land B} \todot \den A \hattimes \den B \\
\end{array}
\]
Observe the eliminations $\andE i$ introduced to reflect conjunctive hypotheses!

Now, we can complete the story for implication.  Let $\tau : \Delta
\leq \Gamma$ and $a : \den A_\Delta$ and set:
\[
\begin{array}{lll}
  \treflect^{A \imp B} & : & \Ne{A \imp B} \todot \den{A \imp B} \\
  \reflect[\Gamma]{A \imp B} t\;\Delta\;\tau\;a & = &
    \reflect[\Delta]B{(\impEP {\Ne[\tau] B t} {(\reify[\Delta] A a)})}
\\[1ex]
  \treify^{A \imp B} & : & \den{A \imp B} \todot \Nf{A \imp B} \\
  \reify[\Gamma]{A \imp B} f & = &
    \impIP {\reify[\Gamma.A]B{(f\;(\weak[A]{\id[\Gamma]})}\;(\reflect[\Gamma.A]A{(\hyp\ttop)})}
\end{array}
\]

\begin{exercise}
  Let $A = ((P \imp Q) \land (Q \imp \top)) \imp Q$ for some atoms
  $P,Q$.  Compute $\reflect[A]A{(\hyp \ttop)}$.
\end{exercise}

\begin{exercise}
  Let $A = (P \imp Q) \imp (P \imp Q)$ for some atoms $P,Q$.
  Compute $\treify^A \comp \treflect^A$.
\end{exercise}

\begin{exercise}
  Using the constructions in this section, show that the category of
  presheaves and natural transformations is cartesian closed.
\end{exercise}

\subsection{Evaluation}

\bibliographystyle{abbrvnat}
\bibliography{auto-nbeSum}

\end{document}