### 2019-03-03 types19/types19.tex
[ types19 ] TYPES abstract

\nonstopmode
\RequirePackage[dvipsnames]{xcolor}

\documentclass[a4paper]{easychair}
\usepackage{amssymb}
%\usepackage{dsmath}
%\usepackage{bbm}
\usepackage{bbold}
\usepackage[utf8x]{inputenc}

\usepackage{natbib}
\usepackage{hyperref}
% \usepackage{agda} % colors

\definecolor{DeepJungleGreen}{rgb}{0.0, 0.29, 0.29}
\definecolor{DarkSpringGreen}{rgb}{0.09, 0.45, 0.27}
\definecolor{ForestGreenTraditional}{rgb}{0.0, 0.27, 0.13}
\definecolor{GreenPigment}{rgb}{0.0, 0.65, 0.31}
\definecolor{CalPolyPomonaGreen}{rgb}{0.12, 0.3, 0.17}
\definecolor{CSSGreen}{rgb}{0.0, 0.5, 0.0}
\definecolor{DarkCerulean}{rgb}{0.03, 0.27, 0.49}
\definecolor{DarkRaspberry}{rgb}{0.53, 0.15, 0.34}

\colorlet{DarkBlue}{MidnightBlue!80!black}
\definecolor{HanBlue}{rgb}{0.27, 0.42, 0.81}
\definecolor{FrenchBlue}{rgb}{0.0, 0.45, 0.73}
\definecolor{EgyptianBlue}{rgb}{0.06, 0.2, 0.65}

\hypersetup{colorlinks=true,
  linkcolor=Red,
  citecolor=DarkRaspberry,%DarkPurple,%DarkCerulean,%CSSGreen,%CalPolyPomonaGreen,%GreenPigment,%ForestGreenTraditional,%DarkSpringGreen,%DeepJungleGreen,%ForestGreen,
  urlcolor=MidnightBlue}

\makeatletter
\renewcommand\bibsection{%
  \section*{\large{\refname}}%
}%
\makeatother

%% Macros

\input{macros}

\renewcommand{\patom}{\ensuremath{o}}
\renewcommand{\tVnf}{\mathsf{Nf}}

\begin{document}

\title{Normalization by Evaluation for Call-by-Push-Value}
\titlerunning{Normalization by Evaluation for Call-by-Push-Value}
\authorrunning{A. Abel}
\author{Andreas Abel}
\institute{
  Department of Computer Science and Engineering,
  Gothenburg University
}

% Keywords:
% Modal typing, linear typing, strictness typing, positivity checking, guarded recursion

\maketitle

%\begin{abstract}
%\end{abstract}

\noindent

Normalization by evaluation (NbE) \citep{bergerSchwichtenberg:lics91}
is the interpretation of
an (open) term of type $A$ as value in a suitable model $\den A$,
followed by \emph{reification} of the value to a normal form of type
$A$.  Functions $f$ in $\den{A \To B}$ are reified as
$\lambda$-abstractions whose bodies are obtained
by \emph{reflecting} a fresh variable of type $A$ as value $a$ in
$\den{A}$ and reifying the application $f\,a$ at type $B$.
A suitable model that supports fresh variable generation
are \emph{presheaves} over the category of
typing contexts $\Gamma$ and order-preserving embeddings
$\Gamma \subseteq \Gamma'$, where a base type $o$ is interpreted as
the presheaf $\tNe\;o$ of neutral normal forms of type $o$, and
function types by the presheaf exponential aka Kripke function space
\citep{catarina:csl93,altenkirchHofmannStreicher:ctcs95}.

NbE for sum types requires a refinement of the model, as reflection of
a variable of type $A + B$ as a value in $\den{A + B}$ requires case
distinction in the model.  One such refinement are \emph{sheaves}
\citep{altenkirchDybjerHofmannScott:lics01}; another is the use of a
monad $\C$
\citep{filinski:semaccounttdpe,filinski:tlca01,barral:PhD}
in the category of presheaves
for the interpretation of sum types:
$\den{A + B} = \C(\den A + \den B)$.
The smallest such \emph{``cover'' monad} $\C$ are binary trees where leaves are
the monadic unit aka $\treturn$, and the nodes case distinctions
over neutrals $\tNe\,(A_1 + A_2)$ of sum type.  When leaves are normal
forms, the whole tree represents a normal form, thus,
$\trunNf : \C(\tNf\,A) \to \tNf\,A$ is trivial.
This \emph{running of the monad}
on normal forms represents the algorithmic part of the sheaf condition
on $\tNf\,A$ and extends as $\trun : \C \den A \to \den A$ to all
semantic types.

The given interpretation of sum types
$\den{A + B} = \C(\den A + \den B)$ corresponds to the call-by-name
(CBN) lambda calculus with lazy constructors.  NbE can also be
performed in call-by-value (CBV) style, then the monad is placed in
the codomain of function types: $\den{A \To B} = \den A \To \C \den B$
\citep{danvy:popl96}.  A systematic semantic analysis of CBN and CBV
lambda-calculi has been pioneered by \citet{moggi:infcomp91} through
translation into his computational lambda calculus;
\citet{filinski:tlca01} studied NbE for the latter calculus using the
continuation monad.  Moggi's work was continued and refined
by \citet{levy:hosc06} who subsumed CBV and CBN under his monadic
call-by-push-value (CBPV) calculus.  In this work, we study NbE for CBPV.

CBPV was designed to study lambda-calculus with effects.  It separates
types into \emph{value} types $P$ and \emph{computation} types $N$, which we, in
analogy to polarized lambda-calculus
\citep{zeilberger:PhD,brockNannestadSchuermann:lpar10,espiritoSanto:entcs17}
refer to as \emph{positive} and \emph{negative} types.
Variables stand for values, thus, have positive types.
The monad that models the effects is placed at the transition from values
to computations $\Comp P$, and computations can be embedded into
values by \emph{thunking} ($\Thunk N$).
\[
\begin{array}{lllrl@{\qquad}l}
  \PTy & \ni & P,Q & ::= & \patom
    % \mid \tyone \mid P_1 \times P_2
    % \mid \tyzero
    \mid P_1 + P_2
    \mid \Thunk N
  & \mbox{positive type / value type}
\\
  \NTy & \ni & N,M & ::= & % \natom
    % \mid \top \mid N_1 \ntimes N_2
    % \mid
    P \To N
    \mid \Comp P
  & \mbox{negative type / computation type}
\\
\end{array}
\]
We restrict to a fragment of \emph{pure} CBPV
with a single positive connective, sum
types $P_1 + P_2$, and a single negative connective, call-by-value
function types $P \To N$.  While we have no proper effects, the
evaluation of open terms requires the effect of case distinction over
neutrals, modeled by a cover monad $\C$.  In the following, we give
inductive definitions of the presheaves of
normal ($\tNf$) and neutral normal forms
($\tNe$) of our fragment of CBPV and a concrete, strong cover monad $\Cov$.
\begin{small}
\begin{gather*}
  \nru{\tvar}{\Var \patom \Gamma}{\PValNf {\patom} \Gamma}
\quad
  \nru{\tthunk}{\Nf N \Gamma}{\PValNfP{\Thunk N}\Gamma}
\quad
  \nru{\tinj_i}{\PValNf{P_i}\Gamma}{\PValNfP{P_1 + P_2}\Gamma}
%\\[2ex]
\quad
  \nru{\tret}
      {\CoverP {\tVNf\,P}  \Gamma}
      {\NfP {\Comp P} \Gamma}
\quad
  \nru{\tabs}
      {\Nf N {(\cext \Gamma P)}}
      {\NfP {P \To N} \Gamma}
\end{gather*}
\end{small}
\begin{small}
\begin{gather*}
  \nru{\tforce}{\VarP {\Thunk N} \Gamma}{\Ne N \Gamma}
\qquad
  \nru{\tapp}
      {\NeP {P \To N} \Gamma \qquad \PValNf P \Gamma}
      {\Ne N \Gamma}
\qquad
  \nru{\tbind}
      {\NeP {\Comp P} \Gamma \qquad \Cover \J (\cext \Gamma P)}
      {\Cover \J \Gamma}
\\[2ex]
  \nru{\treturn}{\J\,\Gamma}{\Cover \J \Gamma}
\qquad
  \nru{\tcase}
      {\VarP {P_1 + P_2} \Gamma
       \qquad \Cover \J (\cext \Gamma {P_1})
       \qquad \Cover \J (\cext \Gamma {P_2})
      }
      {\Cover \J \Gamma}
\end{gather*}
\end{small}
\noindent
($\J$ stands for an arbitrary presheaf in $\Cov\,\J$.)
Normal forms are introductions starting from a variable of base type,
except that the monad is placed at the
transition $\tret$ from positive to negative types ($\Comp P$).
Neutrals are eliminations of variables of type $\Thunk N$ into a
positive type $\Comp P$, and can then be bound (see $\tbind$)
to a variable of type $P$ to be used in a computation.  Variables of
sum type $P_1 + P_2$ can be utilized in computations through a
$\tcase$ split.
\emph{Terms} $\tTm$ of CBPV are obtained by blurring the
distinction between $\tNe$ and $\tNf$,
generalizing $\tbind$ and $\tcase$ from $\Cov\,\J$ to
computations $\tTm\,N$,
relaxing $\tvar$ to variables
of arbitrary type $P$, and $\tforce$ to arbitrary terms of type
$\Thunk N$.  Terms are evaluated in the following presheaf model,
which interprets $\tThunk$ as the identity and $\tComp$ as $\Cov$.
\[
\begin{array}[t]{lcl}
  % \den{\tyone} & = & \hatone \\
  % \den{P_1 \times P_2} & = & \den{P_1} \hattimes \den{P_2} \\
  % \den{\tyzero} & = & \hatzero \\
  \den{P_1 + P_2} & = & \den{P_1} \hatplus \den{P_2} \\
  \den{\Thunk N} & = & \den N \\
  \den{\patom} & = & \tVar\;\patom \\
\end{array}
\qquad
\begin{array}[t]{lcl}
  % \den{\top} & = & \hatone \\
  % \den{N_1 \ntimes N_2} & = & \den{N_1} \hattimes \den{N_2} \\
  % \\
  \den{P \To N} & = & \den P \hatto \den N \\
  \den{\Comp P} & = & \Cov \den P \\
  % \den{\natom}  & = & \C (\tNe\, \natom) \\
\end{array}
\]
The evaluation of $\tbind$-terms in $\Tm N$ relies on
$\trun : \Cov \den N \to \den N$, which makes any computation type
monadic.  Reflection $\uparrow$ and reification $\downarrow$ take the
usual form, only that reflection of positive variables is monadic, to
allow the complete splitting of sums via $\tcase$.
It is invoked by reification of functions $\down {P \To N}$ via $\trunNf$.
\[
\begin{array}[t]{lcl}
  \up P  & : & \tVar\, P \to \Cov\,\den P \\
  \up N & : & \tNe\,N \to \den N \\
\end{array}
\qquad
\begin{array}[t]{lcl}
  \down P & : & \den P \to \tVNf\,P \\
  \down N & : & \den N \to \tNf\,N \\
\end{array}
\]
The details of our construction, plus extension to product types and
polarized lambda calculus, can be found in the full version at
\url{https://arxiv.org/abs/1902.06097}.
A partial Agda formalization is available at
\url{https://github.com/andreasabel/ipl}.

% \[
% \begin{array}[t]{lcl}
%   \upG P  & : & \Var P \Gamma \to \Cov\,\den P \, \Gamma\\
%   \downG P & : & \den P \Gamma \to \VNf P \Gamma \\
% \end{array}
% \qquad
% \begin{array}[t]{l@{\,}l@{~~}c@{~~}l}
%   \multicolumn 2 l {\upG N} & : & \Ne N \Gamma \to \den N \Gamma \\
%   \multicolumn 2 l {\downG N} & : & \den N \Gamma \to \Nf N \Gamma \\
% \end{array}
% \]

% Positive \fbox{$\VNf P \Gamma$} and negative \fbox{$\Nf N \Gamma$}
% normal forms.
% \begin{small}
% \begin{gather*}
%   \nru{\tvar}{\Var \patom \Gamma}{\PValNf {\patom} \Gamma}
% \quad
%   \nru{\tthunk}{\Nf N \Gamma}{\PValNfP{\Thunk N}\Gamma}
% \quad
%   \nru{\tinj_i}{\PValNf{P_i}\Gamma}{\PValNfP{P_1 + P_2}\Gamma}
% %\\[2ex]
% \quad
%   \nru{\tret}
%       {\CoverP {\tVNf\,P}  \Gamma}
%       {\NfP {\Comp P} \Gamma}
% \quad
%   \nru{\tabs}
%       {\Nf N {\cext \Gamma P}}
%       {\NfP {P \To N} \Gamma}
% \end{gather*}
% \end{small}

% Neutral normal forms \fbox{$\Ne N \Gamma$}
% and the strong monad \fbox{$\Cov$}.
% \begin{small}
% \begin{gather*}
%   \nru{\tforce}{\VarP {\Thunk N} \Gamma}{\Ne N \Gamma}
% \qquad
%   \nru{\tapp}
%       {\NeP {P \To N} \Gamma \qquad \PValNf P \Gamma}
%       {\Ne N \Gamma}
% \qquad
%   \nru{\tbind}
%       {\NeP {\Comp P} \Gamma \qquad \Cover \J (\cext \Gamma P)}
%       {\Cover \J \Gamma}
% \\[2ex]
%   \nru{\treturn}{\J\,\Gamma}{\Cover \J \Gamma}
% \qquad
%   \nru{\tcase}
%       {\VarP {P_1 + P_2} \Gamma
%        \qquad \Cover \J (\cext \Gamma {P_1})
%        \qquad \Cover \J (\cext \Gamma {P_2})
%       }
%       {\Cover \J \Gamma}
% \end{gather*}
% \end{small}
% Finally, normal forms of negative types are defined as inductive
% family \fbox{$\Nf N \Gamma$}.  They are generated by maximal negative
% introduction ($\tabs$, $\npair$, $\nunit$) until a negative atom or
% $\Comp P$ is reached.  Then, elimination of neutrals and variables is
% possible through the $\Cov$ monad until an answer can be given in form
% of a base neutral ($\tNe\,\natom$) or a normal value.

% We again can run the cover monad on normal forms, \ie, have
% $\trunNf : \Cov\,(\tNf N) \todot \tNf N$, which extends to
% negative semantic values
% $\trun : \Cov \den N \todot \den N$.

\paragraph*{Acknowledgments.}
%Thanks to the anonymous referees, who helped improving the quality of
%this abstract through their feedback.
This work was supported by
Vetenskapsr\aa{}det under Grant
  No.~621-2014-4864 \emph{Termination Certificates
  for Dependently-Typed Programs and Proofs via Refinement Types}
and by the EU Cost Action CA15123 \emph{Types for programming and verification}.

\begin{footnotesize}

\bibliographystyle{abbrvnat}
\bibliography{auto-types19}

\end{footnotesize}
\end{document}