### 2019-04-01 fscd19/response-fscd19.txt
Minor changes.

A big thanks to the reviewers who spent considerable time to read and
understand our paper and to provide valuable feedback and pointers to
relevant literature.  We will take all comments into account when
preparing the revision.

- Contribution

As the reviewers correctly understand, we are only proving logical
consistency in this paper, via normalization.  Our main contribution
isn't any theorem (consistency of IPL has been proven a gazillion
times before), but the organization of normalization via a monad which
we specify abstractly (cover monad) and for which we give two
implementations (case trees and continuations).

Reviewer 1 voices disappointment that computational soundness is not
demonstrated.  Indeed, we wish we could have included it.  However, we
already struggled to fit the current content into the page limit (as
Reviewer 1 also noticed), and adding more technical content would have
sabotaged the goal of reaching a wide audience through an accessible
presentation (acknowledged by Reviewer 2).  The main idea, the use of
a monad in NbE and its placement at the polarization shift can already
be communicated through the definition of the normalization function
alone, and very effectively, as we believe.

A normalization function with intrinsic soundness is quite a bit more
complicated, see

  https://github.com/andreasabel/ipl/blob/master/src/NbeModel.agda

for the case of IPL, and the added technicalities may cloud the main
idea.  Thus, we save it for a publication venue without page limit
such as a journal.

Without computational soundness, there was no need to retain the
equational theory of IPL, thus, we relegated it to the appendix.
Indeed, our normal forms can be understood purely logically, based on
the distinction of introductions from eliminations and negative
(choiceless) types from positive (choiceful) types.

- Missing theorems (reviewers 1,3)

- Incomplete formalization (reviewers 1,2)

Reviewers 2 and 3 note that the Agda formalization is incomplete and
unpolished.  The Agda formalization helped us to get the constructions
right.  However, we like to think of this paper as an implementation
of Nbe in *Type Theory*.  Systems like Agda disappear (where is LEGO
now?), but type theory is likely to stay with us for a while.  Thus,
we do not consider the Agda code as a contribution of this paper.
Instead, we have strived for a rigorous, self-contained mathematical
presentation.

If wished, we can remove the reference to the partial Agda
formalization from the paper.

- The choice of abstract syntax (Reviewer 3)

Reviewer 3 asks why we do not use the "familiar syntax [of] the
lambda-calculus instead of just the names of the rules?  Could
be some abstract syntax".

In fact we have given an abstract syntax since the rules correspond
1-1 to dependently typed constructors, thus a judgement with rules
corresponds 1-1 to a grammar for elements of an indexed set.  Disposing of
concrete syntax enables us to factor syntax conveniently, for instance
by using Cov in the definition of Nf (line 392).  In case of split₂
(line 438) it is not even immediate what the concrete-syntax analogy
of nesting should be.

- Relation to category theory (reviewers 2,3)

As reviewers 2 and 3 point out, the paper could have been presented
without any reference to category theory.  We use category theory just
as a source of terminology to accurately name the concepts we define.
Further, we like to include experts on category theory in our
readership, who could be upset about not naming the obvious.  The
paper is intended as a bridge between conventional approaches to
lambda-calculus and category-theoretic ones, without demanding too
much expertise.

- CBPV as logical system (Reviewer 2)

In the context of focused lambda-calculi, which we study to reduce the
degrees of liberty in normal forms (in other words, removing
don't-care non-determinism), CBPV arises naturally as calculus with
focusing on positive introductions but not on positive eliminations.
By "coincidence" (or rather not by coincidence, since there are no
coincidences in mathematics) it is the same calculus as the one found
in the search for a lambda-calculus with effects.

We are indeed only interested in CBPV as logical system.  Normalizing effects is
beyond the scope of this paper.  In any case, since effects do not follow the
introduction/elimination scheme, additional considerations would be necessary
for normalizing effects and would like likely depend on the individual effects
considered.

- Future work: relationship to CBN and CBV translations (reviewers 1,2)

We are sorry we could not satisfy the curiousity of our reviewers at
this point.  The idea to investigate the relationship to the CBN/CBV
translations came late in the process of writing, and there was
neither time nor space to address it.

> ----------------------- REVIEW 1 ---------------------
> PAPER: 50
> TITLE: Normalization by Evaluation for Call-by-Push-Value and Polarized Lambda-Calculus
> AUTHORS: Andreas Abel and Christian Sattler
>
> Overall evaluation: 0 (borderline paper)
>
> ----------- Overall evaluation -----------
>
> This paper present a definition of normalization by evaluation for
> call-by-push-value.  The authors first present NbE for a simply type
> lambda calculus with sums. For that, they need to introduce a
> (coverage) monad to be able to do backtracking or to build a case tree
> of possible normalization. A version of this monad is then used to
> interpret computations in CBPV and gives rise to an NbE algorithm for
> CBPV. Finally, the authors discuss modification for focused polarized
> logic, which is closed to CPBV.
>
> General Comment :
>
> This is an interesting work on how to extend NbE technique that are
> compatible with the introduction of effect in the calculus. The paper
> seems mathematically correct and the idea of using such a monad to
> extend NbE to potentially effectful terms is very natural.
>
> However, I have severe reserves about the paper. Therefore I cannot be
> positive about this paper.
>
> - The paper is missing theorem in two senses. First, it is missing a
>   clear statement of what is claimed to be achieved, namely “logical
>   consistency”. The expected theorem are not there.  I was expected at
>   least computational soundness. Also I found quite disappointing that
>   the the combination with the CBN and CBV translation to CBPV has not
>   been investigated, without even explaining why. No time, too
>   difficult, not sure is corresponds to usual NbE ?

To investigate the relationship to the CBN and CBV translations was an
idea we had during the writing, we did not have the time to
investigate it yet.

> - The formalization in Agda i just a link to a GitHub repo that is
>   very hard to dive into. There is no README to know what has been
>   done and where is has been done.  We don’t know what has been
>   formalized in Agda for section 4. It seems that nothing has been
>   formalized for section 3, why ?  The repo looks more like the link
>   to an ongoing project than a link to some documented and finish
>   material aimed at being understood by a reader.

The Agda formalization helped us getting the constructions correct,
but the paper is intended to be self-contained without the Agda
formalization.

> Minor remarks:
>
> The paper is longer than the 15 pages including biblio of the CFP, but
> I didn’t take this into account in my review.
>
> However, putting the equational theory in Appendix rather feels like
> the authors needed to save more space and decided quite arbitrarily to
> cut this.
>
> In the definition of ↑A⇒B ,  where does ∆ comes from ?

When quantifying over (τ : Γ ⊆ Δ) we implicitly quantify over Δ,
similarily as when writing ∀ Δ ⊇ Γ. P, which is a common shorthand for
∀ Δ. Γ ⊆ Δ ⇒ P.

> typos:
> - “a equational” page 5
> - I am not sure "ex falsum quod libet” is correct, one usually says  “ex falso quodlibet”
> - line 307: type of ↓N, P should be N

Thanks, FIXED.

> ----------------------- REVIEW 2 ---------------------
> PAPER: 50
> TITLE: Normalization by Evaluation for Call-by-Push-Value and Polarized Lambda-Calculus
> AUTHORS: Andreas Abel and Christian Sattler
>
> Overall evaluation: 1 (weak accept)
>
> ----------- Overall evaluation -----------
>
> The paper develops normalization by evaluation for simply-typed,
> call-by-push-value and polarized lambda-calculus, all of them
> featuring sums, in a category of presheaves equipped with a
> monad. Besides the fact that the paper seems to give the first
> definition of NbE for CBPV and polarized lambda-calculus, the main
> contribution is the observation how and where to place the monad,
> which simplifies the definitions considerably.
>
> I am not an expert in NbE so I especially enjoyed the very detailed
> and careful introduction of the idea spelled out for the simply-typed
> lambda-calculus. The paper uses categorical terms, but should be
> easily accessible without much categorical knowledge. Every new
> concept is introduced well and category theory is used more as a
> language to describe concrete definitions rather than as a foundation.
>
> NbE for CBPV and polarized lambda-calculus seem to be novel and in my
> belief deserve publication, because it's a standard technique carried
> out for (by now) standard systems in logics and programming languages
> theory. However, I feel that the work is lacking some embedding in
> related work w.r.t the calculi considered. Two or three sentences on
> standard normalization proofs and an explicit sentence that this is
> the first NbE proof for both systems (if it is) would be tremendously
> helpful to judge the value of the contribution.
>
> The treatment of CBPV is more from a logical perspective than as a
> computational system, which is left implicit in the paper. In my
> opinion, the two major aspects of CBPV are (1) the "A value is, a
> computation does"-mantra, which allows fine-grained control over
> reduction and thus enables embedding CBV and CBN and (2) the
> possibility to embed computational effects via the monad. Both aspects
> are not even mentioned, which seems strange to me. Looking at the
> equational theory and not at computation removes the "a value
> is"-aspect, but are there ideas how the NbE method could be adapted to
> also normalize w.r.t call-by-push-value reduction? Can effects be
> easily incorporated into the proof, or would that require completely
> new ideas? Considering that CBPV is in the title, I think this
> questions definitely have to be discussed in the paper.
>
> The paper comes with a partial Agda formalization. I believe the
> paper's claim that this helped in getting the renamings right, but in
> the current state I'm hesitant to count the formalization as valuable
> contribution. Most importantly, the paper needs to make precise what
> is formalized and what is not (possibly in an appendix), where it
> currently just claims a "partial formalization". And then I would hope
> for some explanation why the CBPV part is not formalized. Is it due to
> time pressure? Or would it just be more of the same? If the
> formalization is an explicit contribution I also would expect that
> related work in this area is covered. There must have been formalized
> NbE proofs for simply-typed lambda-calculus before (in Agda and in
> other systems), other formalizations of polarized lambda-calculus and
> of CBPV.
>
>
>
> Comments for the authors
>
> 12: The abstract drops several buzzwords (which I find good), but
> doesn't pick up on them later. For instance, "bi-cartesian" is only
> mentioned in the first sentence of the abstract and never again.
>
> 122: I feel like the choice why the presheaves are all defined on
> Cxt^{op} needs some explanation. The technique I am aware of to build
> categorical models of type theory is to take a category of contexts
> where morally Γ ⊆ Δ gives an arrow Δ → Γ. This confused me more than
> once, especially at points where you said "this is a presheaf"
> omitting the category

In these techniques, one is always looking at a wide subcategory of the category of contexts and substitutions.
An order-preserving embedding Γ ⊆ Δ gives rise to a substitution Δ → Γ in the opposite direction, that's why we have to take presheaves over Cxt^op.
We agree that the direction of arrows in the category Cxt of contexts and order-preserving embeddings is non-standard from this perspective.
However, it has the benefit that our presheaves are just *covariant* functor J : Cxt → Set.

ADDED sentence explaining the relevance of Cxt^op.

> 146: Maybe briefly say what a presheaf is to increase the
> accessibility of the paper even more?

FIXED. Made the introduction of presheaves more intuitive (thought still not
giving a general definition).

> 178: I think it should be stressed even more here that you are talking
> about normal forms up to an equational theory and not up to
> reduction. That's especially important for CBPV, where the
> normal-forms w.r.t. the equational theory completely ignore the
> reduction order induced by call-by-push-value.

Indeed, following the "reduction-free normalization" philosophy of NbE
[3], we are just looking for a canonical representative of the
equivalence classes on terms induced by the equational theory.
However, our normal forms are already fully characterized *logically*
via the bidirectional grammar for Ne/Nf, which emerges from the
introduction/elimination structure and the polarization of the
formulas.  The equational theory matching these normal forms can also
be conceived a *posteriori*, which is the approach we are taking in
this paper.

> 316: I agree that U and F are strange if you do not have lots of
> experience with CBPV. On the other hand, changing the names makes it
> hard for readers which do have experience with CBPV and makes it hard
> for inexperienced readers to match definitions with definitions in
> other papers.

You have a point here, but in any case, we want a common notation for
CBPV and polarized lambda calculus, and a bit longer names (such as
Comp/Thunk) provide more orientation than F/U and the up/down arrows
(which seem very arbitrary; I always confuse them).

> 319: I think talking about CBPV as a fixed definition (as defined in
> the initial paper) rather than a family of definitions is strange. The
> initial CBPV definition is one member of this family, but your
> finitary presentation is also a member. The contrast "CBPV epmloys
> ... while we have ..." is unnecessarily sharp. For instance, there are
> lots of presentation which where the language where a language with
> finitary sums and products is called CBPV, e.g. [1,2].

You are right.  TODO: fix in paper.

> 326: Normally U (F A) is morally the same as A, up to effects. A
> sentence explaining why this is the case or why you deviate here might
> be nice.

This is indeed the case, U.F amounts to the monad, which represents
the effects in our case.

> 361: At least mentioning the standard set-theoretic semantics for CBPV
> with algebras would be good here. There, computations and values map
> to semantically different entities. Do I understand correctly that you
> can use the same entity for both values and computations because
> you're looking at effect-free CBPV?

The difference between positive and negative product does not show up
in the type interpretation but, for instance, in reflection, where we
use `split` in the positive case and the projections in the negative
case.

> 372: Again, I feel some motivation is needed why this are the normal
> forms of CBPV up to the equational theory. When thinking about
> computation, values are always normal because they don't compute, and
> the whole idea of a thunk is that it is a normal form with no
> restrictions.

It is important to remember that we are dealing with open terms and we
are taking eta-equality into account.  Thus, in contrast to the
picture arising from computation (where we only deal with closed terms
and have no eta reduction/expansion), a variable is not always a
normal form, but only if it is of base type or negative type.

> 513: I'm a bit unhappy with the word "implemented" here. Sure, you
> have implemented on paper, but implemented for me means at least
> written down in some formal system/programming language (which is Agda
> here), which is missing for CBPV

Ok, but given the formality of our definitions, the Agda
implementation is a simple exercise.

> 519: What's hard about the agreement proof? Or do you just leave it
> out because you think it's out of scope for the paper?

We are already over the page limit...

> Minor comments
>
> 119: I would maybe write "x : (A ∈ Γ)" to make clear that "x : A ∈ Γ"
> is not a definition here, but you really mean an inhabitant x of a
> type. In named presentations it might be a short hand for "Γ x = A"

WON'T FIX.

>
> 122: Is Cxt a typo? If not, a sentence why Ctx is not used might make
> sense. Is it to avoid confusion with the usual presentation of Ctx
> with arrows inverse to Cxt?

NO COMMENT.

> 146: I would either say "all presheaves will be defined on this and
> this category" or always say which category the presheaf is defined
> on. This makes it way easier for readers to quickly calculate in the
> head that this is really a presheaf.

ADDED. We now say that we only consider presheaves over Ctx^{op}.

> 164: I feel "Needless to say" is unnecessarily snappy to readers with
> only very basic category theory knowledge. As mentioned, your paper is
> doing a really good job in making the presentation accessible, it's a
> good thing that it spells out the details, so I don't think that's
> needless to say at all.

FIXED.

> 178: "an equational theory"

FIXED.

> 201: The symbol for the Kripke function space looks like a hybrid
> between an arrow with a very strange, asymmetric tip and an arrow with
> a hat. I think the latter is intended, so maybe move the hat a little
> bit higher?

FIXED.

> 211: "such an extended context"

FIXED. But was also right before.

> 214: It should be "Ne A Δ"

FIXED. Also A should have been A ⇒ B.

>
> 217: I was confused by the 0 in the index for quite a while. This is a
> singleton context, isn't it? Then maybe write ε.0 or mention that
> singleton contexts look just like types

FIXED.

> 233: I feel repeating the category you build the monad on would
> help. It's on the category of presheaves over Cxt^{op}, right?
>
> 246: There are Γs missing twice on the reifications
>
> 253: the map should be indexed by Δ
>
> 261: again, the map needs a Δ
>
> 270: the Δ is the point your defining on pointwise, but that's too
> implicit currently in my opinion
>
> 315: Two sentences why products appear both negatively and positively
> might be nice for readers not familiar with polarized systems
>
> 319: "is not essential"

FIXED.

> [1] Rizkallah, Christine, Dmitri Garbuzov, and Steve Zdancewic.
> "A Formal Equational Theory for Call-By-Push-Value."
> International Conference on Interactive Theorem Proving. Springer, Cham, 2018.
>
> [2] Forster, Yannick, et al.
> "Call-by-push-value in Coq: operational, equational, and denotational theory."
> Proceedings of the 8th ACM SIGPLAN International Conference on Certified Programs and Proofs. ACM, 2019.
>
> ----------------------- REVIEW 3 ---------------------
> PAPER: 50
> TITLE: Normalization by Evaluation for Call-by-Push-Value and Polarized Lambda-Calculus
> AUTHORS: Andreas Abel and Christian Sattler
>
> Overall evaluation: 1 (weak accept)
>
> ----------- Overall evaluation -----------
>
> Normalization by Evaluation (NbE) is the (by now well-understood)
> beautiful idea that we can normalize terms in a typed lambda-calculus
> through mutually recursive functions for evaluation and reification,
> partially based on types.  NbE is one way to relate programming
> languages (where we don't evaluate under lambda-abstractions) and pure
> proof theory (where we fully normalize).
>
> Sums however present a difficulty.  In Section 2 the present paper
> adapts a prior idea on using monads to update the semantic domain
> needed for NbE to capture sums.  I find this idea compelling and
> NbE factors very cleanly for a lambda-calculus with products and
> sums.
>
> This, however, is characterized as "wasteful" (although I did not
> understand what the authors meant by this), so Section 3 polarizes
> the types and restructures NbE so that the use of the monad is
> confined to the transition from positive to negative types.  This
> seems like a natural idea, clarifying the approach of Section 2.
> The simplicity of the definitions at the end of this section is
> once again compelling, although some complexity is hidden in the
> definition of the semantic functions.
>
> Again, the authors express some dissatisfaction I couldn't quite
> follow and move on to a polarized and focused presentation.  This
> step makes perfect sense to me, although it was difficult for me
> to extract additional insight from this section.
>
> I am not familiar enough with the prior work to judge the degree
> of originality confidently, but the idea of using a monad to
> handle sums in NbE in this manner was new to me, as was the
> extension to the polarized language.  The results have been
> (mostly?) formalized, so I don't have any reason to doubt their
> correctness.  NbE is certainly interesting and I find the results
> to be significant in its further study, even if the open problems
> mentioned in the conclusion are left to future work.
>
> The presentation is idiosyncratic in multiple ways, which one may
> view as either charming or annoying.  Ultimately, I care if ideas
> are communicated well so they can be easily understood, and for me
> the paper receives mixed grades on that score.  I am going into
> some detail on issues of the presentation below.  Overall, I like
> the subject, the paper itself, and I believe I understood the main
> ideas, but I think the presentation can be improved to make it
> more understandable with less background.
>
> Presentation issues:
>
> Intrinsic typing.  Certainly an acceptable choice and seems to
> work well in the formalization.
>
> No syntax.  That presented a significant obstacle to me.  Why not
> use familiar syntax the the lambda-calculus instead of just the
> names of the rules?  Could be some abstract syntax, but even that
> would have helped.

The rules (e.g. for normal forms) are a precise (dependently-typed)
grammar for the abstract syntax (e.g. of normal forms).  The rules are
a pretty-printing of the actual Agda data types that make up the
various (intrinsically typed) syntactic classes.

Note also that restriction to abstract syntax allows for a more
factored presentation of syntax (e.g. Cov in Section 3) and syntax
constructions which are not easily rendered in concrete (linearized)
syntax (such as the nesting of <<P>> in Section 4).

> Presheaves.  In the early parts of the paper there is a lot of
> jargon regarding presheaves, which is, frankly, not necessary to
> understand the paper.

The jargon is indeed not necessary, but it connects to the categorical
notions.  The paper is intended as a bridge between conventional and
categorical meta-theory of lambda-calculus, and we want to be
accessible for both camps.

> No theorems.  The paper presents no formal definitions or theorems
> but simply develops a narrative.  I guess the existence of certain
> dependently typed functions takes the pace of any formal theorems,
> but for me it would still have been helpful to see standard
> theorems regarding the properties of various constructions.  It
> might also help to put some of the definitions into figures so
> they can be referenced in the text (even if the authors want to
> avoid explicit Definition 1, Definition 2, ...).

"No formal definitions" is probably not what was meant here, since, in
fact, *all* definitions in the paper are fully formal (with a few
definitions omitted).
Indeed, as you say, dependently typed total functions second as proofs
of the theorems their types express.  Since we work in Martin-Löf Type
Theory, all are constructions are proof-relevant, thus, there are no
theorems in the original sense.

All constructions we give are uniquely determined by their name, thus,
there is no secondary reference structure such as Definition nnn.  To
help the reader with locating the definition, we have boxed important
definitions which do not have their own disply.

> Other feedback:
>
> The introduction talks about "oracles", but not the body of the
> paper.  I didn't understand what the authors mean here and where
> an "oracle" was supposed to be invoked.

A path in a case tree corresponds to a sequence of questions asked
to the oracle ("is this neutral a left or a right injection?")
with the respective answers.  TODO: explain in paper.

> L.133.  I found it strange to present the types with a polarity
> distinction only at the top level.  P and N are later used (more
> appropriately) in the polarized form.

Strange but convenient. ;-)

> L.136-138.  Why use a backward turnstile?

We flip Γ ⊢ A to A ⊣ Γ so that we can uniformly write the index Γ last
in all presheaves we introduce in the paper.

> L.180-181.  It should be more clear right here what you are
> "proving" in this paper and what is left to future work.  I had
> difficulty interpreting this sentence.

TODO

> L.214. Should this be "Ne A Delta"?

Indeed. FIXED

> L.226. "recapitulate".  Since you do not introduce, discuss, or
> otherwise treat this topic before, "recapitulate" seems like the
> wrong term here.

FIXED. "Recall that"

> L.239.  Actually join seems to be used in a rather essential way
> shortly, so why "uninspiring"?

FIXED formulation.

> L.247.  Not sure what you are trying to say here regarding "pasting".

MOVED to footnote.  It is just to connect to category-theoretic concepts.

> L.249.  "Positive types P are already monadic".  Don't know what
> you mean by that.

[[P]] = C ...

> Section 2.4.  Don't understand the purpose of this section.  Please
> explain.

ADDED an introductory sentence.

> L.336.  "Informally speaking, this makes all negative type monadic".
> I don't understand this sentence.

FIXED.  It means that the effects of monad C are available at any negative type.

> L.384-387.  What is "J" in these rules?  Please explain.

ADDED: J stands for an arbitrary presheaf.

> L.397-405.  I could not understand this paragraph.  For example,
> you talk about "needs to run in a monad".  What does that mean?

It means that ↑ᴾ cannot be a pure function, but must be a monadic function
(see its definition).

> L.422.  "However, the elimination of positive types is still rather
> non-deterministic."  In what sense is it non-deterministic?  What do
> you mean by "It is possible..."  In what sense?

It is non-deterministic in the sense that it is not fully determined,
because "it is possible to only partially break up a composite
positive type and..."