### 2020-07-02 popl21/popl21.tex
[ popl21 ] effect-graded CBPV: typing rules

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{5}
\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2021}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\usepackage{stmaryrd}
\usepackage{xspace}
\usepackage{xypic}

\usepackage{dashrule}
\newcommand{\dashruler}{\hdashrule[0.5ex]{\textwidth}{0.2pt}{1ex}}
\newcommand{\ruler}{\rule{\textwidth}{0.2pt}}

% % https://tex.stackexchange.com/questions/46828/how-to-highlight-important-parts-with-a-gray-background
% \usepackage[breakable, theorems, skins]{tcolorbox}
% \tcbset{enhanced}

% \DeclareRobustCommand{\graytextbox}[2][gray!20]{%
% \begin{tcolorbox}[   %% Adjust the following parameters at will.
%         breakable,
%         left=0pt,
%         right=0pt,
%         top=0pt,
%         bottom=0pt,
%         colback=#1,
%         colframe=#1,
%         width=\dimexpr\textwidth\relax,
%         enlarge left by=0mm,
%         boxsep=5pt,
%         arc=0pt,outer arc=0pt,
%         ]
%         #2
% \end{tcolorbox}
% }

% https://tex.stackexchange.com/questions/502652/define-tcolorbox-in-math-mode
\usepackage{tcolorbox}
\newtcbox{\grayboxtext}{on line,colback=gray!20,colframe=gray!20,size=fbox,arc=0pt,boxrule=0pt}
\newcommand{\graybox}[1]{\grayboxtext{$#1$}}

\newtheorem{remark}{Remark}

\begin{document}

%% Title information
\title[Graded CBPV]{Graded Call-By-Push-Value}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
%\subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{Andreas Abel}
%\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{0000-0003-0420-4492}             %% \orcid is optional
\affiliation{
  \department{Department of Computer Science and Engineering}
  \institution{Gothenburg University}
  \streetaddress{Rännvägen 6b}
  \city{Göteborg}
%  \state{State1}
  \postcode{41296}
  \country{Sweden}
}
\email{andreas.abel@@gu.se}          %% \email is recommended

% %% Author with two affiliations and emails.
% \author{First2 Last2}
% \authornote{with author2 note}          %% \authornote is optional;
%                                         %% can be repeated if necessary
% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
% \affiliation{
%   \position{Position2a}
%   \department{Department2a}             %% \department is recommended
%   \institution{Institution2a}           %% \institution is required
%   \streetaddress{Street2a Address2a}
%   \city{City2a}
%   \state{State2a}
%   \postcode{Post-Code2a}
%   \country{Country2a}                   %% \country is recommended
% }
% \email{first2.last2@inst2a.com}         %% \email is recommended
% \affiliation{
%   \position{Position2b}
%   \department{Department2b}             %% \department is recommended
%   \institution{Institution2b}           %% \institution is required
%   \streetaddress{Street3b Address2b}
%   \city{City2b}
%   \state{State2b}
%   \postcode{Post-Code2b}
%   \country{Country2b}                   %% \country is recommended
% }
% \email{first2.last2@inst2b.org}         %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Call-by-push-value (CBPV) is a simply typed lambda calculus that polarizes
types into value and computation types and can thus express both
call-by-name and call-by-value evaluation in the presence of effects.
Semantically, effects are modeled by a monad, and computation types as
algebras over this monad.
Effect type systems usually express more information than the presence
of an effect; typically effects are structured as pre-ordered monoid
where the monoid operation represents accumulation of effects and the
order subsumption, in analogy to subtyping.
In this work, a \emph{graded} version of is presented where the typing
of computations likens effect typing.  Semantically, computation types
are then represented as graded monad algebras.
Further we present a version of CBPV which additionally has graded coeffects.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
 \begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10003752.10003790.10011740</concept_id>
<concept_desc>Theory of computation~Type theory</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003752.10010124.10010125.10010130</concept_id>
<concept_desc>Theory of computation~Type structures</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003752.10010124.10010138.10010142</concept_id>
<concept_desc>Theory of computation~Program verification</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10003752.10010124.10010131.10010134</concept_id>
<concept_desc>Theory of computation~Operational semantics</concept_desc>
<concept_significance>100</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Theory of computation~Type theory}
\ccsdesc[500]{Theory of computation~Type structures}
\ccsdesc[300]{Theory of computation~Program verification}
\ccsdesc[100]{Theory of computation~Operational semantics}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{effects, coeffects, call-by-name, call-by-value, linear types}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}

Call-by-push-value (CBPV) is a simply typed lambda calculus that polarizes
types into value and computation types and can thus express both
call-by-name and call-by-value evaluation in the presence of effects.
Semantically, effects are modeled by a monad, and computation types as
algebras over this monad.
Effect type systems usually express more information than the presence
of an effect; typically effects are structured as pre-ordered monoid
where the monoid operation represents accumulation of effects and the
order subsumption, in analogy to subtyping.
In this work, a \emph{graded} version of is presented where the typing
of computations likens effect typing.  Semantically, computation types
are then represented as graded monad algebras.
Further we present a version of CBPV which additionally has graded coeffects.

\section{An effect-graded version of CBPV}

%\newcommand{}{\ensuremath{}}
\newcommand{\ie}{\emph{i.e.}}
\newcommand{\eg}{\emph{e.g.}}
\newcommand{\Eg}{\emph{E.g.}}
\newcommand{\loccit}{\emph{loc.\,cit.}}
\newcommand{\den}[1]{\llbracket#1\rrbracket}
\newcommand{\dent}[1]{\llparenthesis#1\rrparenthesis}
\newcommand{\To}{\ensuremath{\Rightarrow}}
\newcommand{\bN}{\ensuremath{\mathbb{N}}}
\newcommand{\tEff}{\ensuremath{\mathsf{Eff}}}
\newcommand{\Eff}{\ensuremath{\mathsf{Eff}}}
\newcommand{\bu}{\ensuremath{\bullet}}
\newcommand{\Ge}{\ensuremath{\varepsilon}}
\newcommand{\GG}{\ensuremath{\Gamma}}
\newcommand{\GS}{\ensuremath{\Sigma}}
\newcommand{\CC}{\ensuremath{\mathcal{C}}}
\newcommand{\SET}{\mathsf{SET}}
\newcommand{\CPO}{\mathsf{CPO}}
\newcommand{\T}[1][]{\mathsf{T}_{#1}}
\newcommand{\trun}{\mathsf{run}}
\newcommand{\run}[1][]{\mathsf{run}_{#1}}
\newcommand{\tA}{\mathsf{A}}
\newcommand{\tB}{\mathsf{B}}
\newcommand{\A}{\mathsf{A}}
\newcommand{\B}{\mathsf{B}}
\newcommand{\tExc}{\mathsf{Exc}}
\newcommand{\Exc}{\mathsf{Exc}}
\newcommand{\String}{\mathsf{String}}
\newcommand{\StringLe}[1]{\mathsf{String}{\leq}\,#1}
\newcommand{\tfmap}{\mathsf{fmap}}
\newcommand{\treturn}{\mathsf{return}}
\newcommand{\tjoin}{\mathsf{join}}
\newcommand{\join}{\mathsf{join}}
\newcommand{\append}{\ensuremath{\mathbin{\mathtt{++}}}}
\newcommand{\tcast}{\mathsf{cast}}
\newcommand{\tid}{\mathsf{id}}

% Types
\newcommand{\tTy}{\mathsf{Ty}}
\newcommand{\Ty}[1]{\tTy^{#1}}
\newcommand{\PTy}{\Ty+}
\newcommand{\NTy}{\Ty-}
\newcommand{\extr}[2]{#1.#2}
\newcommand{\ext}[3]{#1.#2{:}#3}

\newcommand{\sumt}   [2]{\ensuremath{\Sigma_{#1} #2}}
\newcommand{\tupt}   [2]{\ensuremath{\otimes_{#1} #2}}
\newcommand{\rect}   [2]{\ensuremath{\Pi_{#1} #2}}
\newcommand{\sumty}  [3]{\ensuremath{\Sigma_{#1:#2} #3_{#1}}}
\newcommand{\tupty}  [3]{\ensuremath{\otimes_{#1:#2} #3_{#1}}}
\newcommand{\recty}  [3]{\ensuremath{\Pi_{#1:#2} #3_{#1}}}
\newcommand{\thunkty}[2]{\ensuremath{[#1]#2}}
\newcommand{\compty} [1]{\ensuremath{\diamond#1}}

% Terms
\newcommand{\tTm}{\mathsf{Tm}}
\newcommand{\Tm}[1]{\tTm^{#1}}
\newcommand{\PTm}{\Tm+}
\newcommand{\NTm}{\Tm-}
\newcommand{\tthunk}{\mathsf{thunk}}
\newcommand{\thunk}[1]{\tthunk\,#1}
\newcommand{\tforce}{\mathsf{force}}
\newcommand{\force}[1]{\tforce\,#1}
\newcommand{\tin}[1]{\mathsf{in}_{#1}}
\newcommand{\inj}[2]{\tin{#1}\,#2}
\newcommand{\ttup}[1]{\mathsf{tup}^{#1}}
\newcommand{\tup}[2]{\ttup{#1}\,{#2}}
\newcommand{\ptup}[1]{\ttup\,{#1}}
% \newcommand{\ptup}[1]{\ttup{+}\,{#1}}
\newcommand{\ntup}[1]{\ttup{-}\,{#1}}
\newcommand{\tret}{\mathsf{ret}}
%\newcommand{\vlet}[3]{#1\,{=}\,#2;\,#3}
\newcommand{\ret}[1]{\tret\,#1}
\newcommand{\vlet}[3]{#2\,{\mathsf{be}}\,#1.\,#3}
% \newcommand{\let}[3]{#1\,{\leftarrow}\,#2;\,#3}
\newcommand{\bind}[3]{#2\,{\mathsf{to}}\,#1.\,#3}
% \newcommand{\bind}[3]{\ret #1\,{\leftarrow}\,#2;\,#3}
\newcommand{\tsplit}{\mathsf{split}}
\newcommand{\splits}[3]{#1\,{\tsplit}\,#2.\,#3}
\newcommand{\tcases}{\mathsf{cases}}
\newcommand{\case}[2]{#1\,{\tcases}\,\{#2\}}
\newcommand{\caser}[2]{#1\,{\tcases}\,#2}
% \newcommand{\split}[3]{\ptup #1\,{\leftarrow}\,#2;\,#3}
\newcommand{\lam}[2]{\lambda#1.\,#2}
\newcommand{\app}[2]{#1\,#2}
\newcommand{\trecord}{\mathsf{record}}
\newcommand{\recrd}[2]{\trecord_{#1}\,#2}
\newcommand{\record}[2]{\trecord_{#1}\{#2\}}
\newcommand{\tproj}{\mathsf{proj}}
\newcommand{\proj}[2]{\tproj_{#1}\,{#2}}

% Judgements
\newcommand{\ValTy}[3]{#1 \vdash #2 : #3}
\newcommand{\CompTy}[4]{#1 \vdash #2 : #4 \mid #3}
% \newcommand{\CompTy}[4]{#1 \vdash #2 \mathrel{{:}\langle #3 \rangle} #4}
\newcommand{\ru}{\dfrac}
\newcommand{\nru}[3]{#1\;\dfrac{#2}{#3}}
\newcommand{\nrux}[4]{#1\;\dfrac{#2}{#3}\;#4}

\newcommand{\rulename}[1]{\ensuremath{\mbox{\textsc{#1}}}\xspace}
\newcommand{\rintro}[1]{\ensuremath{\mathord{#1}\mbox{-\rulename{intro}}}}
\newcommand{\relim}[1]{\ensuremath{\mathord{#1}\mbox{-\rulename{elim}}}}
\newcommand{\rvar}{\rulename{var}}
\newcommand{\rlet}{\rulename{let}}
\newcommand{\rsub}{\rulename{sub}}

As effect algebra, we assume a pre-ordered monoid
$(\Eff,\bu,\Ge,\leq)$ such that $\_{\bu}\_$ is monotone wrt.\ the
preorder $\leq$ in both arguments.  The unit $\Ge$ shall mean \emph{no
  effect} and the operation $\_{\bu}\_$ serves to accumulate effects,
possibly in a sequential order---unless the monoid is commutative.  The
preorder represents effect subsumption, i.e., loss in precision of the
effect analysis.  Note that the unit $\Ge$ is not necessarily the least element
wrt.\ $\leq$.

\subsection{Recapitulation: modelling effects via monads}

Consider a categorical model $\CC$ of the simply-typed lambda calculus
(STLC), i.e., where types $\tau$ and contexts $\GG$ are interpreted as
objects $\den \tau$ and $\den \Gamma$ of $\CC$ and terms
$\GG \vdash t : \tau$ as morphisms $\dent t \in \CC(\den\GG,\den\tau)$.
Such a category could be $\SET$, interpreting types as sets and terms
as functions, mapping the valuation of their free variables to their
value, or $\CPO$, interpreting types as complete partial orders and
terms as monotone functions, or another model of the STLC.
Typically $\CC$ is a cartesian-closed category, i.e., has products
$A_1 \times A_2$ of objects modelling product types and exponentials $A \To
B$ modeling function types.  Further $\CC$ maybe be distributive,
\ie, have coproducts $A_1 + A_2$ that distribute over products, to
model variants aka disjoint sum types.

Recall that effects are modeled by a suitable monad $\T$ in
$\CC$.  An example would be the \emph{writer monad} $\T\,A = \String \times
A$ that models the effect \emph{output}.
Another example would be the \emph{exception monad} $\Exc + \_$ where
the result of a computation is either an exception $e : \Exc$ or a
regular result.
The monad operations, with
their implementation for writer in $\SET$, are the following:
\[
\begin{array}{lcl}
\tfmap & : & (A \to B) \to \T\,A \to \T\,B \\
\tfmap\, f\, (s,\; a) & = & (s,\, f\, b)
\\[1ex]
\treturn & : & A \to \T\,A \\
\treturn\, a & = & (\verb|""|, a)
\\[1ex]
\tjoin & : & \T\,(\T\,A) \to \T\,A \\
\tjoin\,(s_1,(s_2,a)) & = & (s_1 \append s_2,\; a) \\
\end{array}
\]
Herein, \verb|""| shall denote the empty string and $\append$ string
concatenation.

We assume familiarity with the functor laws $\tfmap\,\tid = \tid$ and
$\tfmap\,(f \circ g) = \tfmap\,f \circ \tfmap\,g$.
The three equational
laws of $\tjoin$ can be visualized compactly in the following
commutative diagram.
\[
\xymatrix@C=12ex{
  \T\,A     \ar[r]^{\treturn} \ar@{=}[dr]^{\tid}
& \T\,(\T\,A) \ar[d]^{\tjoin}
& \T\,(\T\,(\T\,A)) \ar[l]_{\tjoin} \ar[d]^{\tfmap\,\tjoin}
\\
& \T\,A       \ar@{=}[dr]^{\tid}
& \T\,(\T\,A) \ar[l]_{\tjoin}
\\
&
& \T\,A \ar[u]_{\tfmap\,\treturn}
}
\]

\subsection{Effect algebras and graded monads}

We can obtain more information about effects using a monad $\T[e]$
\emph{graded} over elements $e$ of a suitable effect algebra $\Eff$.
Say we want to track an upper bound on the length of the output
produced by a program.  To this end, we can use the preordered monoid
$\Eff = \bN \cup \{\infty\}$ under addition $\_{\bu}\_ = \_{+}\_$
with unit $\Ge = 0$ and the
natural order $\leq$.  The effect $\infty$ then denotes unbounded output, or
output whose length we cannot track in the type system (\eg, when it
depends on some variable).  Increasing the upper bound along $\leq$
means loss of precision of our analysis, with $\infty$ the least
precise information, meaning no upper bound.  The corresponding graded
writer monad is $\T[e]\,A = (\StringLe e) \times A$ where the output is an
element of $\StringLe e$, a string of length at most $e \in \bN \cup
\{\infty\}$.  The operations of a graded monad are, again given with
their $\SET$-implementation for the graded writer:
\[
\begin{array}{lcl}
\tfmap & : & (A \to B) \to \T[e]\,A \to \T[e]\,B \\
\tfmap\, f\, (s,\; a) & = & (s,\, f\, b)
\\[1ex]
\treturn & : & A \to \T[\Ge]\,A \\
\treturn\, a & = & (\verb|""|, a)
\\[1ex]
\tjoin & : & \T[e_1]\,(\T[e_2]\,A) \to \T[e_1 \bu e_2]\,A \\
\tjoin\,(s_1,(s_2,a)) & = & (s_1 \append s_2,\; a)
\\[1ex]
\tcast & : & \T[e] \to \T[e'] \mbox{ for } e \leq e'\\
\tcast & = & \tid \\
\end{array}
\]
A graded version of the exception monad would use effect algebra
$\Eff = \P(\Exc)$ under union and subset; an effect $e$ is a set of
possible exceptions thrown by a computation.\footnote{An example of a
  exception-tracing type system is the Java language.}

The interesting laws for graded monads are given by the following
commutative diagrams.
\[
\xymatrix@C=8ex{
  \T[e]\,A     \ar[r]^{\treturn} \ar@{=}[dr]^{\tid}
& \T[\Ge]\,(\T[e]\,A) \ar[d]^{\tjoin}
\\
& \T[e]\,A
}
\quad
\xymatrix@C=10ex{
  \T[e]\,A     \ar[r]^{\tfmap\,\treturn} \ar@{=}[dr]^{\tid}
& \T[e]\,(\T[\Ge]\,A) \ar[d]^{\tjoin}
\\
& \T[e]\,A
}
\quad
\xymatrix@C=8ex{
  \T[e_1 \bu e_2]\,(\T[e_3]\,A) \ar[d]^{\tjoin}
& \T[e_1]\,(\T[e_2]\,(\T[e_3]\,A)) \ar[l]_{\tjoin} \ar[d]^{\tfmap\,\tjoin}
\\
  \T[e_1 \bu e_2 \bu e_3]\,A
& \T[e_1]\,(\T[e_2 \bu e_3]\,A) \ar[l]_{\tjoin}
}
\]
Further, $\tcast$ commutes with $\tfmap$, namely $\tfmap\,f \circ
\tcast = \tcast \circ \tfmap\,f$, and $\tjoin$, in two ways with
the latter: $\tjoin \circ \tcast = \tcast \circ \tjoin$ and $\tjoin
\circ \tfmap\,\tcast = \tcast \circ \tjoin$.

\subsection{CBPV, monad algebras, and their graded version}

Call-by-push value \cite{levy:hosc06} is a refinement of Moggi's
computational lambda-calculus \cite{moggi:infcomp91} that allows effects
not only in monadic types, \ie, in objects $\T\,A$, but more generally
in computation types.  These correspond to monad algebras for $\T$ in $\CC$,
aka $\T$-algebras.
Those are objects $B$ together with a morphism $\run[B] : \T\,B
\to B$ that allows to formally \emph{run} the monad, ``merging'' its
effects into $B$.
The prime example of a monad algebra is a monadic
type, because $\run[\T\,A] : \T\,(\T\,A) \to \T\,A$ is simply
$\tjoin$.  Levy \citeyearpar{levy:hosc06} shows that monad algebras are
closed under products and exponentiation with arbitrary objects.  \Eg,
in $\SET$ we can define:
\[
\begin{array}{lcl}
%   \run[B_1 \times B_2] & : & \T\,(B_1 \times B_2) \to B_1 \times B_2 \\
%   \run[B_1 \times B_2]\, (s,(b_1,b_2)) & = & (\run[B_1] (s,b_1),\
%                                            \run[B_2] (s,b_2))
% \\[1ex]
  \run[B \times B'] & : & \T\,(B \times B') \to B \times B' \\
  \run[B \times B']\, (s,(b,b')) & = & (\run[B] (s,b),\
                                           \run[B'] (s,b'))
\\[1ex]
  \run[A \To B] & : & \T\,(A \To B) \to (A \To B) \\
  \run[A \To B]\, (s,f) & = & \lambda a.\; \run[B]\,(s,\,f\,a)
\\
\end{array}
\]
These definition implement \emph{lazy} effects that cannot be observed
at computation types such as $A \To B$, but only at value types; the
$\run[]$ of the monad algebra pushes the effects towards result types
that are eventually types of observable objects (values).

With $\run[]$ being a generalization of $\tjoin$, the laws for $\run[]$
are in analogy of those for $\tjoin$:
\[
\xymatrix@C=10ex{
  B     \ar[r]^{\treturn} \ar@{=}[dr]^{\tid}
& \T\,B \ar[d]^{\run[]}
& \T\,(\T\,B) \ar[l]_{\tjoin} \ar[d]^{\tfmap\,\run[]}
\\
& B
& \T\,B \ar[l]_{\run[]}
}
\]

In graded CBPV, monad algebras get replaced by \emph{graded monad
  algebras}.  Given a graded monad $\T$, a $\T$-algebra is a family of
objects $\left(B_e\right)_{e:\Eff}$ and morphisms
$\run[B] : \T[e_1]\,B_{e_2} \to B_{e_1 \bu e_2}$ satisfying these
laws:
\[
\xymatrix@C=10ex{
  B_e     \ar[r]^{\treturn} \ar@{=}[dr]^{\tid}
& \T[\Ge]\,B_e \ar[d]^{\run[]}
\\
& B_e
}
\qquad
\xymatrix@C=10ex{
  \T[e_1 \bu e_2]\,B_{e_3} \ar[d]^{\run[]}
& \T[e_1]\,(\T[e_2]\,B_{e_3}) \ar[l]_{\tjoin} \ar[d]^{\tfmap\,\run[]}
\\
  B_{e_1 \bu e_2 \bu e_3}
& \T[e_1]\,B_{e_2 \bu e_3} \ar[l]_{\run[]}
}
\]
Further the family $B$ should be functorial in the sense that there is
a family of coercion morphisms $B_{e_1 \leq e_2} : B_{e_1} \to B_{e_2}$ with
$B_{e \leq e} = \tid$ and $B_{e_2 \leq e_3} \circ B_{e_1 \leq e_2} =
B_{e_1 \leq e_3}$.

Graded monad algebras are as well closed under pointwise products and
exponentiation with objects, here the graded writer example:
\[
\begin{array}{lcl}
  \run[B \times B'] & : & \T[e_1]\,(B \times B')_{e_2} \to (B \times
                          B')_{e_1 \bu e_2} \\
  \run[B \times B']\, (s,(b,b')) & = & (\run[B] (s,b),\
                                           \run[B'] (s,b'))
\\[1ex]
  \run[A \To B] & : & \T[e_1]\,(A \To B_{e_2}) \to (A \To B_{e_1 \bu e_2}) \\
  \run[A \To B]\, (s,f) & = & \lambda a.\; \run[B]\,(s,\,f\,a)
\\
\end{array}
\]

\begin{figure}[htbp]
\flushleft
\ruler{}
Types.
\[
\begin{array}{lllrl@{\qquad}l}
\PTy & \ni & P
  & ::= & \graybox{\thunkty e N}
     \mid o \mid \sumty i I P \mid \tupty i I P
  & \mbox{Value types (positive types)} \\
\NTy & \ni & N
  & ::= & \compty P
     \mid P \To N \mid \recty i I N
  & \mbox{Computation types (negative types)} \\
\end{array}
\]
\dashruler{}
Terms.
\[
\begin{array}{lllrl@{\qquad}l}
\PTm & \ni & v,w
  & ::= & x
     \mid \thunk t
     \mid \inj i v
     \mid \ptup{\bar v}
  & \mbox{Values (positive terms)} \\
\NTm & \ni & t,u
  & ::= &
         \vlet x v t
  & \mbox{Computations (negative terms):}
\\ &&& \mid &
        \force v
    \mid \case v {\overline{x.\,t}}
    \mid \splits v {\bar x} t
  & \mbox{value eliminations}
\\ &&& \mid &
         \ret v        \mid \bind x u t
  & \mbox{monad operations}
\\ &&& \mid &
         \lam x t      \mid \app t v
  & \mbox{functions}
\\ &&& \mid &
         \trecord\{\overline{i:t}\} % \ntup{\bar t}
       \mid \proj i t
  & \mbox{lazy tuples (records)}
\end{array}
\]
\dashruler{}
Value typing \fbox{$\ValTy \Gamma v P$}\,.
\begin{gather*}
 \nru{\rvar}
     {x{:}P \in \Gamma
    }{\ValTy \Gamma x P}
\qquad
 \nru{\rintro\Box}
     {\CompTy \Gamma t {\graybox e} N
    }{\ValTy \Gamma {\thunk t} {\thunkty {\graybox e} N}}
\\[1.5ex]
 \nru{\rintro\GS}
     {\ValTy \Gamma v {P_i}
    }{\ValTy \Gamma {\inj i v} {\sumt I P}}
\qquad
 \nru{\rintro\otimes}
     {\forall i{:}I, \ \ValTy \Gamma {v_i} {P_i}
    }{\ValTy \Gamma {\ptup v} {\tupt I P}}
    % }{\ValTy \Gamma {\ptup{\overline{v_i}^{i:I}}} {\tupt I P}}
%    }{\ValTy \Gamma {\ptup (v_i)_{i:I}} {\tupt I P}}
\end{gather*}
\dashruler{}
Computation typing \fbox{$\CompTy \Gamma t {\graybox{e}} N$}\,.
\begin{gather*}
 \nru{\rlet}
     {\ValTy \Gamma v P \qquad
      \CompTy {\ext \Gamma x P} t e N
    }{\CompTy \Gamma {\vlet x v t} e N}
\qquad
 \nru{\relim\Box}
     {\ValTy \Gamma v {\thunkty {\graybox e} N}
    }{\CompTy \Gamma {\force v} {\graybox e} N}
\\[1.5ex]
 \nru{\relim\GS}
     {\ValTy \Gamma v {\sumt I P} \qquad
      \forall i{:}I,\ \CompTy {\ext \Gamma {x_i} {P_i}} {t_i} e N
    }{\CompTy \Gamma {\caser v {\{x_i.\, t_i\}_{i:I}}} e N}
\qquad
 \nru{\relim\otimes}
     {\ValTy \Gamma v {\tupt I P} \qquad
      \CompTy {\extr \Gamma {\overline{x_i{:}P_i}^{i:I}}} t e N
    }{\CompTy \Gamma {\splits v {\bar x} t} e N}
\\[1.5ex]
 \nru{\rintro\diamond}
     {\ValTy \Gamma v P
    }{\CompTy \Gamma {\ret v} {\graybox \Ge} {\compty P}}
\qquad
 \nru{\relim\diamond}
     {\CompTy \Gamma u {\graybox{e_1}} {\compty P} \qquad
      \CompTy {\ext \Gamma x P} t {\graybox{e_2}} N
    }{\CompTy \Gamma {\bind x u t} {\graybox{e_1 \bu e_2}} N}
\\[1.5ex]
 \nru{\rintro\To}
     {\CompTy {\ext \Gamma x P} t e N
    }{\CompTy \Gamma {\lam x t} e {P \To N}}
\qquad
 \nru{\relim\To}
     {\CompTy \Gamma t e {P \To N} \qquad
      \ValTy \Gamma v P
    }{\CompTy \Gamma {\app t v} e N}
\\[1.5ex]
 \nru{\rintro\Pi}
     {\forall i{:}I,\ \CompTy \Gamma {t_i} e {N_i}
    }{\CompTy \Gamma {\trecord {\{i : t_i\}_{i:I}}} e {\rect I N}}
    % }{\CompTy \Gamma {\ntup{\overline{t_i}^{i:I}}} e {\rect I N}}
\qquad
 \nru{\relim\Pi}
     {\CompTy \Gamma t e {\rect I N}
    }{\CompTy \Gamma {\proj i t} e  {N_i}}
\\[1.5ex]
\graybox{
 \nrux{\rsub}
      {\CompTy \Gamma t e N}
      {\CompTy \Gamma t {e'} N}
      {e \leq e'}
}
\end{gather*}
\rule{\textwidth}{0.2pt}
  \caption{Effect-graded call-by-push-value.}
  \label{fig:cbpv}
\end{figure}

\subsection{Effect-graded CBPV: syntax and typing}

The syntax and typing rules for effect-graded CBPV are given in
Fig.~\ref{fig:cbpv}.  The differences to pure CBPV are in
\graybox{\mbox{gray boxes}}.

Types are classified into value types $P \in \PTy$ ($A$ in
\citet{levy:hosc06}) and computation types ($\underline B$ in
\loccit).  These are positive and negative types in the terminology of
focusing.  Positive types are generated from base types $o$ via
disjoint sums $\sumty i I P$ with tag set $I$, eager products $\tupty
i I P$ of arity $I$ (composed
from $1$ and $A_1 \times A_2$ in \loccit) and thunking $\thunkty e N$
(written $U\,\underline B$ in \loccit).  In contrast to pure CBPV,
thunk types $\thunkty e N$ are annotated with an effect $e$ that can
be triggered when the thunk is forced.  Negative types are just as in
pure CBPV: monadic types $\compty P$ (written $F\,A$ in \loccit),
function types $P \To N$ (written $A \to \underline B$) in \loccit)
and record types $\recty i I N$ with label set $I$.
Records are \emph{lazy} tuples whose components are only computed by demand.
We abbreviate $\sumty i I P$ by its ``meta-level $\eta$-contraction''
$\sumt I P$; notations $\tupt I P$ and
$\rect I N$ are understood analogously.

Terms are separated into values $v \in \PTm$ and computations $t \in
\NTm$ and identical to pure CBPV, modulo changes in the concrete
syntax.  Values introduce positive types,
computations introduce
negative types and
eliminate both positive and negative types.
We use bars to indicate sequences, e.g., $\bar v$ for a
sequence of values, but drop the bar when the context of discourse
makes clear that we are dealing with sequences rather than single
objects.  For instance, ``$v_i$'' (where $i : I$) in the premise of
\rintro\otimes indicates that $v$ is a sequence of values with
elements $(v_i)_{i:I}$.  We may abbreviate $\trecord \{i :
t_i\}_{i:I}$ by $\recrd I t$ where $I$ is the label set and $t$ a
mapping from labels $i:I$ to terms $t_i$.

The meaning of the term constructors is best understood via their
typing.  Typing contexts $\Gamma$ are finite maps from variables $x$
to value types $P$, with $\ext \Gamma x P$ denoting the update of the
finite map $\Gamma$ at key $x$ with value $P$.

Value typing $\ValTy \Gamma v P$ is just as in pure CBPV, however,
computation typing $\CompTy \Gamma t e N$ also records effects $e :
\Eff$ potentially produced at runtime by computation $t$.  Thunking a
computation (rule \rintro\Box) stores the inferred effect classifier
$e$ in the thunk type $\thunkty e N$.

Effects are accumulated via the introduction and elimination rule for
the graded monad.  The unit $\ret v$ of the monad is effect-free
(\rintro\diamond), running this computation just produces the pure
value $v$.  Sequencing computations $u$ and $t$ via the \emph{bind}
construct $\bind x u t$ composes the effects $e_1$ of $u$ with the
effects $e_2$ of $t$ in that order.  The intuition is that first $u$
is run, producing effects classified by $e_1$, and its result is bound
to $x$ to run $t$, producing effects classified by $e_2$.  The sum of
the effects is classified by $e_1 \bu e_2$.

The other introduction and elimination rules are just as in pure
CBPV, except that they propagate the effect classifier $e$ from
hypothesis to conclusion.  Note that in case distinction (\relim\GS)
and record construction (\rintro\Pi) all subterms $t_i$ are required
to produce effects classified by the same $e$.  However, in reality,
different branches of a case distinction, for instance, may produce
very different effects.  To end up with a unique classifier $e$, the
branches may have to be typed using effect subsumption (\rsub).  In
fact, the uses of \rsub can be confined to the hypotheses of \relim\GS
and \rintro\Pi, except for a final invokation of \rsub at the very end
of the typing derivation.  Alternatively, we could have introduced
effect algebras with suprema $\sup_{i:I} e_i$ instead of a preorder $e
\leq e'$.  However, suprema might not always exist; by using
subsumption \rsub, we delegate the problem of partiality to the
construction of a typing derivation.

\begin{remark}\label{rem:cbpv}
  We recover pure CBPV from graded CBPV using the trivial effect
  algebra $\Eff = \{\Ge\}$.
\end{remark}

\begin{example}

\end{example}

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.

  This material is based upon work supported by the
  Swedish Research Council (Vetenskapsrådet)
  under Grant
  No.~2019-04216 \emph{Modal Dependent Type Theory}.

  % This material is based upon work supported by the
  % \grantsponsor{GS100000001}{National Science
  %   Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  % No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  % No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  % conclusions or recommendations expressed in this material are those
  % of the author and do not necessarily reflect the views of the
  % National Science Foundation.
\end{acks}

%% Bibliography
\bibliography{medium}

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

\end{document}